A modern, event-based architecture for distributed evolutionary algorithms

### 2018-04-03 paper.tex
Shorter
But now Figure 3 is alone in a page. Will continue to reduce the length
and worry about it later. @JJ Please check (when you have time) if I am
not missing something crucial.

\section{Introduction}
Cloud computing is increasingly becoming the dominant way of running
the server side of most enterprise applications nowadays, the same as the browser is
the standard platform for the client side. Besides the convenience of the pay-as-you-go
model, it also offers a way of describing the infrastructure as part
of the code, so that it is much easier to reproduce results and has 
been a boon for scientific computing.
%The availability of resources only limited by your budget, and in many
%cases for free, as well as the inherent reproducibility of cloud
%deployments, has been a boon for scientific computing. 
However,
programming the cloud means that monolithic applications, that is,
applications built on a single stack of services that communicate by
layers, are no longer an efficient architectural design for scientific workflows. Cloud
architectures favor asynchronous communication over heterogeneous %favor: american
resources, and shifting from mostly sequential and monolithic 
to an asynchronously parallel architecture will also imply important
reformulation of the algorithms in order to take full advantage of these technologies.

  Cloud native applications add a layer of abstraction to the
  underlying distributed computing system, seamlessly integrating
  different elements in a single data flow, allowing the user to just focus on code and service connections. Services are {\em native} points
  in this new architecture, departing from a monolithic or even
  distributed paradigm to become a loosely collection of
services, in fact {\em microservices} \cite{microservices}, which in
 many cases are stateless, reacting to some event and
  {\em living} only while they are doing some kind of processing.
%   and not keeping state from one invocation to the next.
% redundant - Mario  
  Reactive systems not only allow massive scaling and independent deployment they are also more economical than other monolithic options.  Platform as a service (PaaS) or even Container as a Service (CaaS) approaches need to be running all
  the time in order to maintain their state, so they are paid for their size and
  time they remain active.

  At any rate, while one of the main selling points of Functions as a Service (FaaS) is their ultra-fast activation time, from our point of view their most interesting feature is the fact that they provide stateless processing. An important caveat
  of stateless processing is that algorithms must be adapted to this fact and
  turned, at least in part, into a series of stateless steps thaworking on a data stream.

    It is also taken to an atomic extreme
    with the so-called serverless architectures \cite{Varghese2018849},
    which allow vendors and users to deploy code as single, stateless
    functions, that get activated via {\em rules}, {\em triggers} or
    explicitly, reacting to events consumed from
    a message queue.
    %or bus that
    % generally reacting to events from a message queue or bus that
    % what is generated? rules are triggered and triggers fire in
    % response to events
    % Please change it to whatever you think it's the best - JJ
    %is the backbone of the application. 
    The first commercial
    implementation of this kind of architecture was released by Amazon
    with its Lambda product, to be closely followed by releases by Azure
    (Microsoft) and Google and OpenWhisk, an open source implementation
    released by IBM \cite{Baldini2016287}.

  These architectures are {\em event-based} in the sense that program
  flow is governed by events.
  %, which in this case are going to be simply
  %messages received by the MessageHub service.
  %These {\em events} trigger actions,
  %and actions create new message/events which start the loop all over
  %again. This is not the only mode of action in general serverless
  %architectures; however, since they define a temporal axis and make new
  %actions depend on the result of other actions, is the one that we are
  %going to choose in this paper to implement a stateless version of an
  %evolutionary algorithm.

  There is usually no direct translation of most algorithms to this
  architecture. This is specially true in evolutionary algorithms, who
  rely on state, in the shape of a population, to perform their search
  in the space of solutions. It is relatively straightforward to convert
  an evolutionary algorithm to a stateless version by simply making a
  {\sf generation} function $g$ that takes a population as a function
  and returns another population. A simple serverless GA would then be
  triggered by the creation of a initial population, and would perform
  one or several generations in a function, returning the result to the
  message hub and thus triggering a new action. This would be
  functionally equivalent to a traditional, serial evolutionary
  algorithm, and might be cheaper than a XaaS (where X can be P as in
  platform, I as in infrastructure, C as in container or even DB, since
  whole systems can be implemented using a database engine as a server framework) 
  implementation since it
  would eschew some overhead. However, it would also incur in some
  overhead by reading and writing to the message queue and, since it
  would be working on a single population, no scaling could be done,
  since there would be a single pipeline.

  In this paper we want to introduce KafkEO, a serverless framework for
  evolutionary algorithms and other population-based systems. The main
  design objective is to leverage the scaling capabilities of a
  serverless framework, as well as create a system that can be deployed
  on different platforms by using free software. Our intention has
  also been to create an algorithm that is functionally equivalent to an asynchronous
  parallel, island-based, EA, which can use parallelism and at the same
  time reproduce mechanisms that are akin to migration. The island-based 
  paradigm is relatively standard in distributed EA applications, but in our case, 
  we have been using it since it allows for better parallelism
  and thus performance, at the same time it makes keeping diversity
  easier while needing fewer parameters to tune.

  %%% Maybe a little more about the why Island?
  %%% Not only better performance in time, but also, more diverstity and less
  %%% parametrization
  % Please add something and I'll review it. - JJ
  % OK, TO DO - M
  % And I have done it now... - JJ
  We will examine the results of this framework on some BBOB benchmarks,
  taking into account
  %speed as well as
  the number of evaluations, comparing it
  against another Cloud ready parallel pool based implementation (with
  containers).
  %The speed measurements are intended mainly as baseline,
  %since no effort has been made to optimize it.
  The
  implementation is also free software and can be downloaded from
  GitHub.

  % \begin{table}
  %   \caption{Frequency of Special Characters}
  %   \label{tab:freq}
  %   \begin{tabular}{ccl}
  %     \toprule
  %     Non-English or Math&Frequency&Comments\\
  %     \midrule
  %     \O & 1 in 1,000& For Swedish names\\
  %     $\pi$ & 1 in 5& Common in math\\
  %     \$ & 4 in 5 & Used in business\\
  %     $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  %   \bottomrule
  % \end{tabular}
  % \end{table}

>>>>

\section{Introduction}
Cloud computing is increasingly becoming the dominant way of running
the server side of most enterprise applications nowadays, the same as the browser is
the standard platform for the client side. Besides the convenience of the pay-as-you-go
model, it also offers a way of describing the infrastructure as part
of the code, so that it is much easier to reproduce results and has 
been a boon for scientific computing.
%The availability of resources only limited by your budget, and in many
%cases for free, as well as the inherent reproducibility of cloud
%deployments, has been a boon for scientific computing. 
However,
programming the cloud means that monolithic applications, that is,
applications built on a single stack of services that communicate by
layers, are no longer an efficient architectural design for scientific workflows. Cloud
architectures favor asynchronous communication over heterogeneous %favor: american
resources, and shifting from mostly sequential and monolithic 
to an asynchronously parallel architecture will also imply important
reformulation of the algorithms in order to take full advantage of these technologies.
  Cloud native applications add a layer of abstraction to the
  underlying distributed computing system, seamlessly integrating
  different elements in a single data flow, allowing the user to just focus on code and service connections. Services are {\em native} points
  in this new architecture, departing from a monolithic or even
  distributed paradigm to become a loosely collection of
services, in fact {\em microservices} \cite{microservices}, which in
 many cases are stateless, reacting to some event and
  {\em living} only while they are doing some kind of processing.
%   and not keeping state from one invocation to the next.
% redundant - Mario  
  Reactive systems not only allow massive scaling and independent deployment they are also more economical than other monolithic options.  Platform as a service (PaaS) or even Container as a Service (CaaS) approaches need to be running all
  the time in order to maintain their state, so they are paid for their size and
  time they remain active.
  At any rate, while one of the main selling points of Functions as a Service (FaaS) is their ultra-fast activation time, from our point of view their most interesting feature is the fact that they provide stateless processing. An important caveat
  of stateless processing is that algorithms must be adapted to this fact and
  turned, at least in part, into a series of stateless steps thaworking on a data stream.
    It is also taken to an atomic extreme
    with the so-called serverless architectures \cite{Varghese2018849},
    which allow vendors and users to deploy code as single, stateless
    functions, that get activated via {\em rules}, {\em triggers} or
    explicitly, reacting to events consumed from
    a message queue.
    %or bus that
    % generally reacting to events from a message queue or bus that
    % what is generated? rules are triggered and triggers fire in
    % response to events
    % Please change it to whatever you think it's the best - JJ
    %is the backbone of the application. 
    The first commercial
    implementation of this kind of architecture was released by Amazon
    with its Lambda product, to be closely followed by releases by Azure
    (Microsoft) and Google and OpenWhisk, an open source implementation
    released by IBM \cite{Baldini2016287}.

  %These architectures are {\em event-based} in the sense that program
  %flow is governed by events.
  %, which in this case are going to be simply
  %messages received by the MessageHub service.
  %These {\em events} trigger actions,
  %and actions create new message/events which start the loop all over
  %again. This is not the only mode of action in general serverless
  %architectures; however, since they define a temporal axis and make new
  %actions depend on the result of other actions, is the one that we are
  %going to choose in this paper to implement a stateless version of an
  %evolutionary algorithm.

  %There is usually no direct translation of most algorithms to this
  %architecture. This is specially true in evolutionary algorithms, who
  %rely on state, in the shape of a population, to perform their search
  %in the space of solutions. 
  %It is relatively straightforward to convert
  %an evolutionary algorithm to a stateless version by simply making a
  %{\sf generation} function $g$ that takes a population as a function
  %and returns another population. A simple serverless GA would then be
  %triggered by the creation of a initial population, and would perform
  %one or several generations in a function, returning the result to the
  %message hub and thus triggering a new action. This would be
  %functionally equivalent to a traditional, serial evolutionary
  %algorithm, and might be cheaper than a XaaS (where X can be P as in
  %platform, I as in infrastructure, C as in container or even DB, since
  %whole systems can be implemented using a database engine as a server framework) 
  %implementation since it
  %would eschew some overhead. However, it would also incur in some
  %overhead by reading and writing to the message queue and, since it
  %would be working on a single population, no scaling could be done,
  %since there would be a single pipeline.

  There is usually no direct translation of most algorithms to this
  architecture. This is specially true in evolutionary algorithms, who
  rely on state, in the shape of a population, to perform their search
  in the space of solutions.
  In this paper we want to introduce KafkEO, a serverless framework for
  evolutionary algorithms and other population-based systems. The main
  design objective is to leverage the scaling capabilities of a
  serverless framework, as well as create a system that can be deployed
  on different platforms by using free software. Our intention has
  also been to create an algorithm that is functionally equivalent to an asynchronous
  parallel, island-based, EA, which can use parallelism and at the same
  time reproduce mechanisms that are akin to migration. The island-based 
  paradigm is relatively standard in distributed EA applications, but in our case, 
  we have been using it since it allows for better parallelism
  and thus performance, at the same time it makes keeping diversity
  easier while needing fewer parameters to tune.

  %%% Maybe a little more about the why Island?
  %%% Not only better performance in time, but also, more diverstity and less
  %%% parametrization
  % Please add something and I'll review it. - JJ
  % OK, TO DO - M
  % And I have done it now... - JJ

  %% TODO: 
  %% Explain a little more BBOB the benchmarks
  %% - Mario

  We will examine the results of this framework on some BBOB benchmarks,
  taking into account
  %speed as well as
  the number of evaluations, comparing it
  against another Cloud ready parallel pool based implementation (with
  containers).
  %The speed measurements are intended mainly as baseline,
  %since no effort has been made to optimize it.
  The
  implementation is also free software and can be downloaded from
  GitHub.
  %% TODO: Add link?
